### 2007-01-09 Documentation/Paper2/paper2.tex
JK - Added paragraph about numerical efficiency in the paper

>>>>

Sophisticated \GLOBES\ simulations with several experiments and complicated
parameter correlations can take several days to finish even on modern computers.
To mitigate this problem as far as possible, we have taken several steps to
optimize the performance of the multi-dimensional minimizations, which are at
the heart of each \GLOBES\ $\chi^2$ analysis.

Firstly, the \GLOBES\ minimizers distinguish between minimization over
oscillation parameters and minimization over systematics parameters.
This is necessary because each step in the oscillation parameter space
requires a time-consuming re-computation of oscillation probabilities,
while a modification of the systematics parameters does not. The standard
procedure is to perform a full minimization in systematics space after each
step in oscillation space. This avoids interference of the two categories
of parameters, and is known to have excellent convergence properties.
The numerical method that does the minimization is the Powell
algorithm~\cite{Press:NumRecip}.

In \GLOBES\ 3.0, we provide an alternative method, which does the full minimization
in one go and is therefore several times faster in most cases. In this algorithm,
each Powell iteration over oscillation parameters is followed by one iteration
over systematics parameters. Since there may be some pathological cases where this
interleaved minimization strategy behaves differently than the old method,
it is by default not used in \GLOBES\ 3.0. This ensures maximal compatibility
with old application programs, but it requires users who wish to benefit from the
new algorithm to explicitly select it by an API function call.

Besides optimizing the minimization functions, it is also mandatory to
optimize the routines which are called by them to calculate $\chi^2$
for one specific set of parameters. In particular, the computation of
oscillation probabilities, which has to be done every time the oscillation
parameters have changed, should be as efficient as possible. The most expensive
step here is the diagonalization of the hermitian $3\times3$ neutrino Hamilton
operator in matter. The LAPACK rotuines~\cite{Anderson:LAPACK}, which were
used to solve this eigenproblem in previous versions of \GLOBES, turned out
not to be the optimal choice because they are mainly optimized for very large
matrices~\cite{Kopp:2006wp}. In the case of the small and usually well-conditioned
matrices appearing in \GLOBES, they spend too much time evaluating their parameter
list and planning the optimal diagonalization strategy. Therefore, \GLOBES\ now
relies an algorithm which has been designed specifically for the diagonalization
of hermitian $3\times3$ matrices. It is essentially the extremely fast analytical
algorithm discussed in~\cite{Kopp:2006wp}, but for the rare cases where this may become
inaccurate, it falls back to the QL method from~\cite{Press:NumRecip}.
